Thermodynamic response functions in a cell fluid model

### 2024-09-01 main_morse_response_functions.tex
write two equivalent expressions for beta_V suitable for two forms of eos

	\subsection{Thermal pressure coefficient}
	The thermal pressure coefficient is defined by
	\begin{equation}
		\label{def:therm_pres_coef}
		\beta_V = \left( \frac{\partial P}{\partial T} \right)_{V,\langle N \rangle}.
	\end{equation}
	We rewrite $\beta_V$ into a form that is suitable for the equation of state~\eqref{eq:eosMT}
	\begin{eqnarray}
		\beta_V & = & \left(\frac{\partial P}{\partial T}\right)_{V, N}
		\nonumber\\
		& = & \left(\frac{\partial P}{\partial T}\right)_{\mu} + \left(\frac{\partial P}{\partial \mu}\right)_T \left(\frac{\partial \mu}{\partial T}\right)_{V, N}.
	\end{eqnarray}
	Applying the cyclic relation
	\begin{equation*}
		\left(\frac{\partial \mu}{\partial T}\right)_{V, N} 
		\left(\frac{\partial T}{\partial V}\right)_{\mu, N}
		\left(\frac{\partial V}{\partial \mu}\right)_{T, N}
		= -1
	\end{equation*}
	we obtain
	\begin{eqnarray*}
		\left(\frac{\partial \mu}{\partial T}\right)_{V, N} 
		& = & 
		- \left(\frac{\partial V}{\partial T}\right)_{\mu, N} 
		\left(\frac{\partial V}{\partial \mu}\right)^{-1}_{T, N} \\
		& = & - \left(\frac{\partial \bar{n}}{\partial T}\right)_{\mu}
		\left(\frac{\partial \bar{n}}{\partial \mu}\right)^{-1}_{T}.		
	\end{eqnarray*}
	Substituting this result into the formula for $\beta_V$, we arrive at the final expression for the thermal pressure coefficient
	\begin{equation}
		\beta_V = \left(\frac{\partial P}{\partial T}\right)_{\mu} 
		- \left(\frac{\partial P}{\partial \mu}\right)_T 
		\left(\frac{\partial \bar{n}}{\partial T}\right)_{\mu}
		\left(\frac{\partial \bar{n}}{\partial \mu}\right)^{-1}_{T}
	\end{equation}
	which is easily calculated based on Eqs.~\eqref{eq:eosMT} and~\eqref{eq:density}.\\
	(NB. Compare this equation with Eq.(17) from~\cite{StrokerMeier2021})

	We introduce the reduced thermal pressure coefficient by
	\begin{equation}
		\beta^*_V = \frac{v}{k_{\rm B}}\beta_V.
	\end{equation}

>>>>

	\subsection{Thermal pressure coefficient}
	The thermal pressure coefficient is defined by
	\begin{equation}
		\label{def:therm_pres_coef}
		\beta_V = \left( \frac{\partial P}{\partial T} \right)_{V,\langle N \rangle}.
	\end{equation}
	We rewrite $\beta_V$ into a form that is suitable for the equation of state $P=P(T,\mu)$, see~\eqref{eq:eosMT} and~\eqref{eq:eosPTM_reduced}
	\begin{eqnarray}
		\beta_V & = & \left(\frac{\partial P}{\partial T}\right)_{V, N}
		\nonumber\\
		& = & \left(\frac{\partial P}{\partial T}\right)_{\mu} + \left(\frac{\partial P}{\partial \mu}\right)_T \left(\frac{\partial \mu}{\partial T}\right)_{V, N}.
	\end{eqnarray}
	Applying the cyclic relation
	\begin{equation*}
		\left(\frac{\partial \mu}{\partial T}\right)_{V, N} 
		\left(\frac{\partial T}{\partial V}\right)_{\mu, N}
		\left(\frac{\partial V}{\partial \mu}\right)_{T, N}
		= -1
	\end{equation*}
	we obtain
	\begin{eqnarray*}
		\left(\frac{\partial \mu}{\partial T}\right)_{V, N} 
		& = & 
		- \left(\frac{\partial V}{\partial T}\right)_{\mu, N} 
		\left(\frac{\partial V}{\partial \mu}\right)^{-1}_{T, N} \\
		& = & - \left(\frac{\partial \bar{n}}{\partial T}\right)_{\mu}
		\left(\frac{\partial \bar{n}}{\partial \mu}\right)^{-1}_{T}.		
	\end{eqnarray*}
	Substituting this result into the formula for $\beta_V$, we arrive at the final expression for the thermal pressure coefficient
	\begin{equation}
		\beta_V = \left(\frac{\partial P}{\partial T}\right)_{\mu} 
		- \left(\frac{\partial P}{\partial \mu}\right)_T 
		\left(\frac{\partial \bar{n}}{\partial T}\right)_{\mu}
		\left(\frac{\partial \bar{n}}{\partial \mu}\right)^{-1}_{T}
	\end{equation}
	which is easily calculated based on Eqs.~\eqref{eq:eosMT} and~\eqref{eq:density}.\\
	(NB. Compare this equation with Eq.(17) from~\cite{StrokerMeier2021})

	We introduce the reduced thermal pressure coefficient by
	\begin{equation}
		\beta^*_V = \frac{v}{k_{\rm B}}\beta_V.
	\end{equation}
	It is expressed in terms of the reduced quantities $P^*$ and $\bar{n}$ as follows
	\begin{equation}
		\beta^*_V = \varepsilon\beta_c 
		\left[ \left(\frac{\partial P^*}{\partial \tau}\right)_{\mu} 
		- \left(\frac{\partial P^*}{\partial \mu}\right)_T 
		\left(\frac{\partial \bar{n}}{\partial \tau}\right)_{\mu}
		\left(\frac{\partial \bar{n}}{\partial \mu}\right)^{-1}_{T} 
		\right].
	\end{equation}

	If it is preferable to use the equation of state in the form $P=P(T, \bar{n})$, see~\eqref{eq:eosNT} and~\eqref{eq:eosPTn_reduced}, then the most suitable expressions for the thermal pressure coefficient follow immediately from the definition~\eqref{def:therm_pres_coef}
	\begin{equation}
		\beta_V = \left(\frac{\partial P}{\partial T}\right)_\rho
	\end{equation}
	and
	\begin{equation}
		\beta^*_V = \varepsilon\beta_c \left(\frac{\partial P^*}{\partial \tau}\right)_{\rho}
	\end{equation}
